%% thesis.tex 2014/04/11
%
% Based on sample files of unknown authorship.
%
% The Current Maintainer of this work is Paul Vojta.

% For a masters thesis, replace the above \documentclass line with
% \documentclass[masters]{ucbthesis}
% This affects the title and approval pages, which by default calls this
% document a "dissertation", not a "thesis".

\documentclass{ucbthesis}

% To compile this file, run "latex thesis", then "biber thesis"
% (or "bibtex thesis", if the output from latex asks for that instead),
% and then "latex thesis" (without the quotes in each case).

% Double spacing, if you want it.  Do not use for the final copy.
% \def\dsp{\def\baselinestretch{2.0}\large\normalsize}
% \dsp

% If the Grad. Division insists that the first paragraph of a section
% be indented (like the others), then include this line:
% \usepackage{indentfirst}

\input{_preamble}

\begin{document}

\input{configuration}

\maketitle
% Delete (or comment out) the \approvalpage line for the final version.
% \approvalpage
\copyrightpage

\include{abstract}

\begin{frontmatter}

\begin{dedication}
\null\vfil
\begin{center}
\vspace{12pt}

To my parents, whose gentle support made it possible to pursue this challenging
journey.

\vspace{22pt}

To my brother, Aria, who enriched and challenged my style of thinking with
stimulating conversations, and whose work ethic and intellectual curiosity have
been an inspiration.

\vspace{22pt}

To the great friends whose excellent company ensured that the last five years
were marked by myriad memorable moments, from convivial conversations to
a month-long trip abroad.

\end{center}
\vfil\null
\end{dedication}

% You can delete the \clearpage lines if you don't want these to start on
% separate pages.

\tableofcontents
\clearpage
\listoffigures
\clearpage
\listoftables

\begin{acknowledgements}

The period marked by the conclusion of this work has been shaped by several
remarkably skilled and dedicated mentors. First and foremost, I wish to thank
Mark van der Laan for being an unrivaled advisor and an inspirational
collaborator. Mark's outstanding guidance, tireless support, and incredible
patience made this work possible, and his creativity and generosity made working
with him a great pleasure. I hope our years working closely together have
instilled in me the fundamentals of Mark's remarkably well-structured approach
to problem solving. I owe my first experiences in applied biostatistics research
to Alan Hubbard, who gave me the opportunity to explore the intricate
intersection of applied statistics, computational biology, and bioinformatics.
Nick Jewell spent his time in teaching me about survival analysis and its
fascinating applications, and shared with me his seasoned perspective on our
field and its history.

I have benefited considerably from the efforts of researchers in the earlier
stages of their careers too. During the first year of my graduate studies,
I luckily crossed paths with David Benkeser, whose work on statistical methods
in vaccine trials captivated me. I quickly found in David a dedicated and
patient mentor to whom I owe many of the skills and much of the knowledge that
I rely upon daily. In turn, David embraced my then-nascent interests in
reproducible research and statistical software design.
%Our work appears in the second, third, and fifth chapters of this dissertation.
A few years later, I met Iv\'{a}n D\'{i}az, whose emerging interests in causal
mediation analysis led to an unpredictably fruitful collaboration. I have
learned a great deal from watching Iv\'{a}n's exemplary skills in causal
inference and statistical theory in action.
%Parts of our work formed the fourth and fifth chapters of this dissertation.
I eagerly look forward to postdoctoral training and longstanding collaborations
with David and Iv\'{a}n both.

Over the years, conversations and courses with Fernando P{\'e}rez, Jarrod
Millman, Philip Stark, and Sandrine Dudoit sparked the embers of my interests in
developing open source software for statistics and in the need for standards of
reproducible research in the computational sciences. Their influential examples
showed me that these important ideas could have major impacts on the practice of
statistics and the transparency of modern science. These important ideas would
become cornerstones of my personal research philosophy and greater research
agenda.

A few truly amazing individuals --- close colleagues and friends --- stand out
across the recent years. I have grown immensely from numerous deep and
interesting conversations with Rachael Phillips, Ivana Malenica, and Jeremy
Coyle. Together, we built the earliest version of the \texttt{tlverse} software
ecosystem (\url{https://github.com/tlverse}), delivered well-received workshops,
and nearly completed a book. Our work has been richly rewarding, and,
better still, we even found some time for fun along the way. I will recall our
adventures fondly.

\end{acknowledgements}

\end{frontmatter}

\pagestyle{headings}

\include{chap1}
\include{chap2}
\include{chap3}
\include{chap4}
\include{chap5}
\include{chap6}

\printbibliography

\end{document}
